---
title: "Medical Appoinment No Show"
output: 
  prettydoc::html_pretty:
    fig_height: 4
    highlight: pygments
    theme: cayman
    toc: yes
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Wrangling

### Load packages

```{r load-packages, message = FALSE}
library(tidyverse)
```

### Read table and Cleaning data

```{r read table, message=FALSE, warning=FALSE}
# read the table

df <-  read_csv("data/KaggleV2-May-2016.csv")

```

```{r check data, message=FALSE, warning=FALSE}
# check the data

head(df)
```

```{r check structure of the data, message=FALSE, warning=FALSE}
# check structure of the data
str(df)
```

```{r change columns name}

# change columns name
names(df)<-  c('patient_id',
                   'appointment_id',
                   'gender','schedule_day',
                   'appointment_day',
                   'age',
                   'neighborhood',
                   'scholarship',
                   'hypertension',
                   'diabetes',
                   'alcoholism',
                   'handicap',
                   'sms_received',
                   'no_show')
```

```{r check again}

# check again
head(df)
```


```{r check missing value}
# check missing value

sapply(df,function(x)sum(is.na(x)))
```

```{r  change data type of some columns}
# change data type of some columns
df <- mutate_at(df, vars('gender',
                   'neighborhood',
                   'scholarship',
                   'hypertension',
                   'diabetes',
                   'alcoholism',
                   'handicap',
                   'sms_received'), as.factor)
```

```{r check data type}
# check data type
lapply(df, class)
```


```{r  check summary statistics}
# check summary statistics
summary(df)
```

minimum age is −1, so lets check all the value less than 0

```{r check value less than 0}
#check value less than 0
df[df$age < 0, ]

```

```{r drop the row with condition}
#drop the row with condition

df <-df[!(df$age<0),]
```

```{r}
summary(df$age)
```


## Exploratory Analysis



### Age vs No Show

#### Is there any relation between Age and No show?

#### plot and statistics

In the dataset description, it mention that in no_show column No means patient showed up and yes means patient did not showed up

```{r}
# age vs showed up or not
# blue showed up
ggplot(df,aes(x=age)) + 
    geom_histogram(data=subset(df,no_show == 'No'),fill = '#00BFC4', alpha = 0.8) +
    geom_histogram(data=subset(df,no_show == 'Yes'),fill = '#F8766D', alpha = 0.8) 
    
```

Since there is confusion with no and yes with showed up or not, lets make it clear

```{r}
# in no_show column No means patient showed up and yes means patient did not showed up
df$no_show[df$no_show == 'No'] <- 'Showed up'
df$no_show[df$no_show == 'Yes'] <- 'not showed up'
```

```{r}
df$no_show <- as.factor(df$no_show)
```


```{r age box plot}

# age box plot

ggplot(df, aes(x = no_show, y = age, fill = no_show))+
    geom_boxplot()
```


```{r statistics}
#statistics

select(df, age, no_show) %>% 
    group_by(no_show) %>% 
    summarise(age_mean = mean(age))
```

#### more visualization

```{r fig.height=8, fig.width=16}
ggplot(data = df)+
    geom_bar(aes(factor(age), fill = no_show), position = position_fill())+
    theme(axis.text.x = element_text(size= 12, angle = 90, hjust = 1))
```

#### t test 

```{r make separate column}

# make column of age

showed_up <- df[df$no_show == 'Showed up',]$age
not_showed_up <- df[df$no_show == 'not showed up',]$age
```

```{r age t test}
# independent two sample t test
t.test(showed_up, not_showed_up)
```

```{r other way t test}
# just easy way, same result
t.test(df$age ~ df$no_show)
```

### Gender vs No Show

#### Is there any relation between Gender and No Show?

#### Plot and statistics

```{r replace levels}
# replace levels
levels(df$gender)[levels(df$gender)=="M"] <- "Male"
levels(df$gender)[levels(df$gender)=="F"] <- "Female"

```


```{r plot gender}
# count gender
ggplot(data = df)+
    geom_bar(aes(x = gender,fill = gender))
```



```{r gender table}
# gender table
table(df$gender, df$no_show)
```

```{r}
# total patient
table(df$no_show)
```



```{r gender vs noshow stacked bar}
#gender vs noshow stacked bar
ggplot(df)+
    geom_bar(aes(x = gender, fill = no_show))
```

```{r proportion table}
# proportion table
prop.table(table(df$gender, df$no_show), margin = 1)
```


```{r plot proportionally}
#plot of proportion data, just add position_fill()
ggplot(df)+
    geom_bar(aes(x = gender, fill = no_show), position = position_fill())
```


#### chi-squared test

```{r}
chisq.test(table(df$gender,df$no_show))
```
without continuity correction

```{r}
chisq.test(table(df$gender,df$no_show), correct = FALSE)
```

### Scholarship, Hypertension, Diabetes, Alcohol, Handicap, Sms_received vs No Show

#### What are the relation of No Show with Scholarship, Hypertension, Diabetes, Alcohol, Handicap, Sms_received?

#### plot

```{r eval=FALSE, include=FALSE}
# below code does create plot but not good looking . could not figure out how to loop using ggplot
df_plot <- select(df, scholarship, 
               hypertension,
               diabetes, 
               alcoholism, 
               handicap, 
               sms_received)

column_plot <- colnames(df_plot)
par(mfrow = c(4,2))

for(i in 1:ncol(df_plot)){
    counts <- table(df_plot[,i])
    barplot(counts)
}
    




```



```{r multiple columns plot: count, fig.height=12, fig.width=8}

# Count patients number in each group
library(gridExtra)

p1 <- ggplot(df)+geom_bar(aes(scholarship, fill = scholarship))
p2 <- ggplot(df)+geom_bar(aes(hypertension, fill =hypertension))
p3 <- ggplot(df)+geom_bar(aes(diabetes, fill = diabetes))
p4 <- ggplot(df)+geom_bar(aes (alcoholism, fill =  alcoholism))
p5 <- ggplot(df)+geom_bar(aes( handicap, fill = handicap))
p6 <- ggplot(df)+geom_bar(aes(sms_received, fill = sms_received))

grid.arrange(p1,p2,p3,p4,p5,p6, nrow = 3)




```

```{r multiple columns plot vs noshow ,fig.height=12, fig.width=8}
# this plot for each group with patient who showed up and not showed up
p1 <- ggplot(df)+geom_bar(aes(scholarship, fill = no_show))
p2 <- ggplot(df)+geom_bar(aes(hypertension, fill =no_show))
p3 <- ggplot(df)+geom_bar(aes(diabetes, fill = no_show))
p4 <- ggplot(df)+geom_bar(aes (alcoholism, fill =  no_show))
p5 <- ggplot(df)+geom_bar(aes( handicap, fill = no_show))
p6 <- ggplot(df)+geom_bar(aes(sms_received, fill = no_show))

grid.arrange(p1,p2,p3,p4,p5,p6, nrow = 3)
```

```{r multiple columns plot: proportion, fig.height=12, fig.width=8}

# this plot is same as above one execpt it show proportion.
p1 <- ggplot(df)+geom_bar(aes(scholarship, fill = no_show), position = position_fill())
p2 <- ggplot(df)+geom_bar(aes(hypertension, fill =no_show), position = position_fill())
p3 <- ggplot(df)+geom_bar(aes(diabetes, fill = no_show), position = position_fill())
p4 <- ggplot(df)+geom_bar(aes (alcoholism, fill =  no_show), position = position_fill())
p5 <- ggplot(df)+geom_bar(aes( handicap, fill = no_show), position = position_fill())
p6 <- ggplot(df)+geom_bar(aes(sms_received, fill = no_show), position = position_fill())

grid.arrange(p1,p2,p3,p4,p5,p6, nrow = 3)
```



#### chi-squared test 

```{r chi-squared test multiple column, message=FALSE, warning=FALSE}
# chi-squared test

chisq.test(table(df$scholarship, df$no_show), correct = FALSE) # scholarship
chisq.test(table(df$hypertension, df$no_show), correct = FALSE) # hypertension
chisq.test(table(df$diabetes, df$no_show), correct = FALSE) # diabetes
chisq.test(table(df$alcoholism, df$no_show), correct = FALSE) #alcoholism
chisq.test(table(df$handicap, df$no_show), correct = FALSE) #handicap
chisq.test(table(df$sms_received, df$no_show), correct = FALSE) # sms_received

```



### Appointment day vs No Show

####plot

```{r appointment day vs noshow}
# make new columns 
df$day <- weekdays(as.Date(df$appointment_day))
```


```{r}
ggplot(df)+geom_bar(aes(day, fill = day))
```

```{r}
library(lubridate)

# make days column, with lebel true
df$date <- as.Date(df$appointment_day)
df$days <- wday(df$date, label=TRUE)
```

```{r plot}

# day start monday and goes on by day, using days column
ggplot(df)+geom_bar(aes(days, fill = day))
```

```{r days vs no show}

# days vs no show
ggplot(df)+geom_bar(aes(days, fill = no_show))
```

```{r}
#days vs no show, using days column
ggplot(df)+geom_bar(aes(days, fill = no_show), position = position_fill())
```


#### chi-squared test

used day column because days column contain sunday.

```{r}
# day column
table(df$day, df$no_show)
```

```{r}
#days column
table(df$days, df$no_show)
```


```{r}
chisq.test(table(df$no_show,df$day))
```

### Neighborhood vs No Show



```{r}
df_neighbor <- data.frame(table(df$neighborhood, df$no_show))
names(df_neighbor) <- c("neighborhood", "no_show", 'Count')
head(df_neighbor)
```

#### plot

```{r fig.height=8, fig.width=16}

ggplot(df_neighbor)+
geom_bar(aes(x = reorder(neighborhood, -Count), y = Count, fill = no_show), stat = 'identity')+
    theme(axis.text.x = element_text(size= 12, angle = 90, hjust = 1))



```



```{r fig.height=8, fig.width=16}

# proportion 
ggplot(df)+
    geom_bar(aes(x = neighborhood, fill = no_show), position = position_fill())+
    theme(axis.text.x = element_text(size= 12, angle = 90, hjust = 1))

```

## Logistic Regression

```{r}
df_log <- select(df, age, gender, scholarship, hypertension, diabetes, handicap,sms_received, no_show)
```

```{r}
model <- glm(no_show ~ . , data = df_log ,family = "binomial")
```

```{r}
summary(model)
```

```{r}
df_log <- select(df, age, scholarship, hypertension, diabetes,sms_received, no_show)

model <- glm(no_show ~ . , data = df_log ,family = "binomial")
summary(model)
```

